---
title: "Network topology from circuit motifs"
---

```{r Rmdsetup, echo=FALSE}
# Set max width of markdown code outputs
options(width = 10000)
```

The simplest mathematical models of neural networks are built from homogeneous McCulloch-Pitts neurons and scalar-weight connections between them. Biological neural networks, such as those in patches of cortex, are far more complex. They contain different types of neurons each with their own electrical behavior and spatially extended connections. Computational neuroscientists standardly model biological neural networks as [dynamical systems](https://philpapers.org/rec/ELIMBM-3), including the [growth-transform](https://doi.org/10.3389/fnins.2020.00425) (GT) framework used in this package. Instead of a simple one-dimensional weight between two homogeneous neurons, the GT framework makes use of both a transconductance parameter and a temporal modulation factor to determine network behavior. Roughly, the transconductance parameter is the inverse of a traditional weight, while the temporal modulation factor allows for capturing the different electrodynamics of different neuron types. 

<h2>Load spike rasters</h2>

Set up the R environment by clearing the workspace, setting a random-number generator seed, and loading the neurons package.

```{r Rsetup}
# Clear the R workspace to start fresh
rm(list = ls())

# Set seed for reproducibility
set.seed(12345) 

# Load neurons package
library(neurons) 
```

All 

```{r load_rasters}
cortical.patch <- new.network()
plot.network(cortical.patch)


# ...



rm(list = ls())
set.seed(12345) 
library(neurons) 

rf <- list()
for (i in 1:5) {
  rf[[i]] <- matrix(0.5, nrow = 1, ncol = 1)
}

cortical.patch <- new.network()
cortical.patch <- set.network.structure(
  cortical.patch,
  layer_names = paste0("L", 1:5),
  n_layers = 5,
  n_columns = 5,
  neurons_per_node = matrix(10, nrow = 5, ncol = 1),
  recurrence_factors = rf
  )
plot.network(cortical.patch)


```

