---
title: "Estimating network time constants using dichotomized Gaussians"
---

```{r Rmdsetup, echo=FALSE}
# Set max width of markdown code outputs
options(width = 10000)
```

Intro here.

<h2>Load spike rasters</h2>

The first step is to set up the R environment by clearing the workspace, setting a random-number generator seed, and loading the neurons package.

```{r Rsetup}
# Clear the R workspace to start fresh
rm(list = ls())

# Set seed for reproducibility
set.seed(12345) 

# Load neurons package
library(neurons) 
```

Rasters from patch-clamp data.

```{r load rasters}
spike.rasters.exper <- read.csv(
  system.file(
    "extdata", 
    "spike_rasters_2022data.csv", 
    package = "neurons"
    )
  )
```

<h3>Making neurons</h3>

With the data preprocessed into spike rasters, the next step is to convert them into a special class of object from the neuron package, **neuron**. 

```{r create_neuron_objects}
neurons.exper <- load.rasters.as.neurons(spike.rasters.exper)
```

We want to load the spike rasters as **neuron**s because **neuron**s come with built-in methods for plotting rasters, plotting autocorrelation, and estimating autocorrelation parameters with dichotomized Gaussian simulations. 
